Safe Transferable Regions

### 2018-03-22 paper/ecoop18_rebuttal.txt

4. As far as the syntax is concerned, `transfer` and `free` are like
any other methods, hence their application is modeled by the generic
method application expression ($e.m<...>(...)$). Since the operational
semantics models a single actor (and not a system of actors), the
behavior of `free` is same as that of `transfer`: both result in the
region becoming unavailable to the actor calling `transfer`/`free`.
The rule for `transfer` captures this semantics.

5. The run-time checks are precisely captured by the premises of
the opsem rules that deal with exceptions (namely, OpenTransferred and
TransferOpened). The checks basically check whether the action being
performed (e.g., `transfer`) is allowed given the current typestate of
the region (e.g., "live"), raising an exception otherwise.

6. `inRgn.free()` in Fig. 3 is needed to prevent a memory leak.
Reference going out of scope doesn't automatically free a region,
because the region may have been aliased. Naiad code does contain
`free` statements.

7. The immediate difference between Kilim and Broom seems to be the
enforcement of linearity of references to transferable objects. Kilim
relies on ownership and borrowing techniques (ala Rust) to ensure that
transferable objects are not aliased. Broom on the other had lets
transferable regions to be aliased freely, allowing idiomatic C# code
(e.g., iterators) to work without modifications.

### 2018-05-29 paper/ecoop18_rebuttal.txt

Review #50C
-----------

1. How the type system differentiates between `free` and `transfer`:
The type system doesn't. Dynamic regions in Broom are primarily
intended to transfer data among actors (hence the name "Transferable
Regions"). It is therefore expected that a dynamic region is
transferred at least once during its lifetime (before it is freed),
thereby making it unsafe to allow outgoing references from the region
to other regions.  While nothing prevents a transferable region to be
used merely as dynamic memory, that is not an intended usecase for
which Broom is optimized.

2. Serializing: Broom's region runtime [7] serializes data if the
receiver of a transferable region is remote.

3. Our model does admit aliases (through the `let` expression). The
operational semantics defines a *store* (Σ) that maps locations to
their typestates, and implicitly models a heap at the granularity of a
region (each location $l$ denotes a region, and every object whose
first region parameter is $l$ (e.g., `new B<l,..>)(...)`) is allocated
in the region at $l$). Mutable object fields constitute an ostensible
omission from our model (also the FGJ model), but this isn't a major
shortcoming considering that constructor application (during the new
object creation) includes assignments to object fields. We would
include a detailed note to this extent in the final version.

### 2018-05-29 paper/implementation.tex
Edits.

Next, we translated 4 out of the 6 Naiad streaming query operator
benchmarks (Naiad vertices) used in~\cite{Broom:HotOS} to
$\absof{\FB^+}$, and used \namec to verify their safety (The remaining
two Naiad vertices are not interesting from the type system
perspective).  The number of LOC performing operations on \C{Region}
objects relative to the total LOC is 8\% or under in the Naiad
benchmarks.  During the process, we found multiple instances of
potential memory safety violations in the $\absof{FB^+}$ translation
of all the 4 Naiad vertices, which we verified to be present in the
original C\# implementation as well. The cause of all safety
violations is the creation of a reference from the outgoing message (a
transferable region) to the payload of the incoming message. For
example, the implementation of \C{SelectVertex} contains the
following:
\begin{codejava}
  if (this.selector(inMsg.payload[i])) {
    outMsg.set(outputOffset, inMsg.payload[i]);
    ...
  } 
\end{codejava}
The \C{outMsg} is later transferred to a downstream actor, where the
reference to \C{inMsg}'s payload becomes unsafe\footnote{This unsafe
reference could have gone unnoticed during experiments
in~\cite{Broom:HotOS} because their experimental setup included only
one actor.}. We eliminated such unsafe references by creating a clone
of \C{inMsg.payload[i]} in \C{outMsg}, and our compiler was
subsequently able to certify the safety of all references.

>>>>

Next, we translated 4 out of the 6 Naiad streaming query operator
benchmarks (Naiad vertices) used in~\cite{Broom:HotOS} to
$\absof{\FB^+}$, and used \namec to verify their safety. The 2
remaining benchmarks were left out because their region behavior (from
the perspective of the type system) is subsumed by the included
benchmarks. The number of LOC performing operations on \C{Region}
objects relative to the total LOC is 8\% or under in the Naiad
benchmarks.  During the process, we found multiple instances of
potential memory safety violations in the $\absof{FB^+}$ translation
of all the 4 Naiad vertices, which we verified to be present in the
original C\# implementation as well. The cause of all safety
violations is the creation of a reference from the outgoing message (a
transferable region) to the payload of the incoming message. For
example, the implementation of \C{SelectVertex} contains the
following:
\begin{codejava}
  if (this.selector(inMsg.payload[i])) {
    outMsg.set(outputOffset, inMsg.payload[i]);
    ...
  } 
\end{codejava}
The \C{outMsg} is later transferred to a downstream actor, where the
reference to \C{inMsg}'s payload becomes unsafe\footnote{This unsafe
reference could have gone unnoticed during experiments
in~\cite{Broom:HotOS} because their experimental setup included only
one actor.}. We eliminated such unsafe references by creating a clone
of \C{inMsg.payload[i]} in \C{outMsg}, and our compiler was
subsequently able to certify the safety of all references.
